---
title: "Get BLUPs combining all trial data"
author: "Marnin Wolfe"
date: "2021-May-03"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, 
                      tidy='styler', tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
```

# Previous step 

1. [Prepare training dataset](01-cleanTPdata.html): Download data from DB, "Clean" and format DB data.

# Get multi-trial BLUPs from raw data (two-stage)

Two-stage procedure: 

1. Fit mixed-model to multi-trial dataset and extract BLUPs, de-regressed BLUPs and weights. Include two rounds of outlier removal.
2. Genomic prediction with drg-BLUPs from multi-trial analysis as input. 

**Work below represents Stage 1 of the Two-stage procedure.**

## Set-up training datasets
```{bash, eval=F}
# activate multithread OpenBLAS for fast matrix algebra
export OMP_NUM_THREADS=56
```
```{r, message=F}
rm(list=ls())
library(tidyverse); library(magrittr);
source(here::here("code","gsFunctions.R"))
dbdata<-readRDS(here::here("output","NRCRI_ExptDesignsDetected_2021May03.rds"))
traits<-c("CGM","CGMS1","CGMS2","MCMDS",
          "DM","DMsg","PLTHT","BRNHT1","HI",
          "logDYLD","logFYLD","logTOPYLD","logRTNO")

# **Nest by trait.** Need to restructure the data from per-trial by regrouping by trait. 
dbdata<-nestDesignsDetectedByTraits(dbdata,traits)
```

```{r}
dbdata %>% mutate(N_blups=map_dbl(MultiTrialTraitData,nrow)) %>% rmarkdown::paged_table()
```

To fit the mixed-model I used last year, I am again resorting to `asreml`. I fit random effects for rep and block only where complete and incomplete blocks, respectively are indicated in the trial design variables. `sommer` should be able to fit the same model via the `at()` function, _but_ I am having trouble with it _and_ `sommer` is much slower even without a dense covariance (i.e. a kinship), compared to `lme4::lmer()` or `asreml()`. 

```{r}
dbdata %<>% 
  mutate(fixedFormula=ifelse(Trait %in% c("logDYLD","logFYLD","logRTNO","logTOPYLD"),
                             "Value ~ yearInLoc + PropNOHAV","Value ~ yearInLoc"),
         randFormula=paste0("~idv(GID) + idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                            "+ at(IncompleteBlocks,'Yes'):blockInRep"))
dbdata %>% 
  mutate(Nobs=map_dbl(MultiTrialTraitData,nrow)) %>% 
  select(Trait,Nobs,fixedFormula,randFormula) %>% 
  rmarkdown::paged_table()
```

```{r}
# randFormula<-paste0("~vs(GID) + vs(trialInLocYr) + vs(at(CompleteBlocks,'Yes'),repInTrial) + vs(at(IncompleteBlocks,'Yes'),blockInRep)")
# library(sommer)
# fit <- mmer(fixed = Value ~ 1 + yearInLoc,
#             random = as.formula(randFormula),
#             data=trainingdata,
#             getPEV=TRUE)
```

## Function to run asreml

Includes rounds of outlier removal and re-fitting.

```{r, eval=F}
fitASfunc<-function(fixedFormula,randFormula,MultiTrialTraitData,...){
  # test arguments for function
  # ----------------------
  # MultiTrialTraitData<-dbdata$MultiTrialTraitData[[7]]
  # #Trait<-dbdata$Trait[[3]]
  # fixedFormula<-dbdata$fixedFormula[[7]]
  # randFormula<-dbdata$randFormula[[7]]
  # test<-fitASfunc(fixedFormula,randFormula,MultiTrialTraitData)
  # ----------------------
  require(asreml); 
  fixedFormula<-as.formula(fixedFormula)
  randFormula<-as.formula(randFormula)
  # fit asreml 
  out<-asreml(fixed = fixedFormula,
              random = randFormula,
              data = MultiTrialTraitData, 
              maxiter = 40, workspace=800e6, na.method.X = "omit")
  #### extract residuals - Round 1
  
  outliers1<-which(abs(scale(out$residuals))>3.3)
  
  if(length(outliers1)>0){
    
    x<-MultiTrialTraitData[-outliers1,]
    # re-fit
    out<-asreml(fixed = fixedFormula,
                random = randFormula,
                data = x, 
                maxiter = 40, workspace=800e6, na.method.X = "omit")
    #### extract residuals - Round 2
    outliers2<-which(abs(scale(out$residuals))>3.3)
    if(length(outliers2)>0){
      #### remove outliers
      x<-x[-outliers2,]
      # final re-fit
      out<-asreml(fixed = fixedFormula,
                  random = randFormula,
                  data = x, maxiter = 40,workspace=800e6, na.method.X = "omit")
    }
  }
  if(length(outliers1)==0){ outliers1<-NULL }
  if(length(outliers2)==0){ outliers2<-NULL }
  
  ll<-summary(out,all=T)$loglik
  varcomp<-summary(out,all=T)$varcomp
  Vg<-varcomp["GID!GID.var","component"]
  Ve<-varcomp["R!variance","component"]
  H2=Vg/(Vg+Ve)
  blups<-summary(out,all=T)$coef.random %>%
    as.data.frame %>%
    rownames_to_column(var = "GID") %>%
    dplyr::select(GID,solution,`std error`) %>%
    filter(grepl("GID",GID)) %>%
    rename(BLUP=solution) %>%
    mutate(GID=gsub("GID_","",GID),
           PEV=`std error`^2, # asreml specific
           REL=1-(PEV/Vg), # Reliability
           drgBLUP=BLUP/REL, # deregressed BLUP
           WT=(1-H2)/((0.1 + (1-REL)/REL)*H2)) # weight for use in Stage 2
  out<-tibble(loglik=ll,Vg,Ve,H2,
              blups=list(blups),
              varcomp=list(varcomp),
              outliers1=list(outliers1),
              outliers2=list(outliers2))
  return(out) }
```

## Run asreml
```{r, eval=F}
library(furrr); options(mc.cores=14); plan(multiprocess)
library(asreml)
dbdata %<>% 
  mutate(fitAS=future_pmap(.,fitASfunc))
dbdata %<>%
  select(-fixedFormula,-randFormula,-MultiTrialTraitData) %>%
  unnest(fitAS)
```

## Output file
```{r, eval=F}
saveRDS(dbdata,file=here::here("output","NRCRI_blupsForModelTraining_twostage_asreml_2021May03.rds"))
```

# Results

See [Results](05-Results.html): Home for plots and summary tables.

# Next step

3. [Check prediction accuracy](03-CrossValidation.html): Evaluate prediction accuracy with cross-validation.

