---
title: "Genomic predictions"
author: "Marnin Wolfe"
date: "2021-May-03"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, 
                      tidy='styler', tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
```

# Previous step

3. [Check prediction accuracy](03-CrossValidation.html): Evaluate prediction accuracy with cross-validation.

# Objective

**Current Step**  

4. [Genomic prediction](04-GetGBLUPs.html): Predict _genomic_ BLUPs (GEBV and GETGV) for all selection candidates using all available data.

# Set-up
```{bash, eval=F}
cd /home/jj332_cas/marnin/NRCRI_2021GS/; 
export OMP_NUM_THREADS=56 # activate multithread OpenBLAS 
```

```{r, eval=F}
library(tidyverse); library(magrittr); 
source(here::here("code","gsFunctions.R"))
A<-readRDS(file=here::here("output","Kinship_A_NRCRI_2021May03.rds"))

# BLUPs from the 2 stage procedure
# (stage 1 of 2) 
blups<-readRDS(file=here::here("output","NRCRI_blupsForModelTraining_twostage_asreml_2021May03.rds")) %>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-`std error`) %>% 
  filter(GID %in% rownames(A)) %>% 
  nest(TrainingData=-Trait)
```

# Prediction

`runGenomicPredictions()`

cbsurobbins (112 cores; 512GB) 

Model A
```{r, eval=F}
options(future.globals.maxSize= 1500*1024^2)
predModelA<-runGenomicPredictions(blups,modelType="A",grms=list(A=A),gid="GID",ncores=14)
saveRDS(predModelA,file = here::here("output","genomicPredictions_ModelA_twostage_NRCRI_2021May03.rds"))
```

Model ADE
```{r, eval=F}
D<-readRDS(file=here::here("output","Kinship_D_NRCRI_2021May03.rds"))
AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2021May03.rds"))
options(future.globals.maxSize= 3000*1024^2)
predModelADE<-runGenomicPredictions(blups,modelType="ADE",grms=list(A=A,D=D,AD=AD),gid="GID",ncores=14)
saveRDS(predModelADE,file = here::here("output","genomicPredictions_ModelADE_twostage_NRCRI_2021May03.rds"))
```

# Write GEBV/GETV to disk

```{r}
rm(list=ls()); gc()
library(tidyverse); library(magrittr);
predModelA<-readRDS(file = here::here("output","genomicPredictions_ModelA_twostage_NRCRI_2021May03.rds"))
predModelADE<-readRDS(file = here::here("output","genomicPredictions_ModelADE_twostage_NRCRI_2021May03.rds"))
traits<-c("CGM","CGMS1","CGMS2","MCMDS",
          "DM","DMsg","PLTHT","BRNHT1","HI",
          "logDYLD","logFYLD","logTOPYLD","logRTNO")
```

```{r}
unique_gids<-predModelA %>%
  dplyr::select(genomicPredOut) %>%
  unnest(genomicPredOut) %>%
  select(-varcomps) %>%
  unnest(gblups) %$%
  GID %>%
  unique

c1a<-unique_gids %>%
  grep("c1a",.,value = T,ignore.case = T) %>%
  union(.,unique_gids %>%
          grep("^F",.,value = T,ignore.case = T) %>%
          grep("c1b",.,value = T,ignore.case = T,invert = T))
c1b<-unique_gids%>% grep("c1b",.,value = T,ignore.case = T)
c2a<-unique_gids %>%
  grep("C2a",.,value = T,ignore.case = T) %>%
  grep("NR17",.,value = T,ignore.case = T)
c2b<-unique_gids %>%
  grep("C2b",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a)]
c3a<-unique_gids %>%
  grep("C3a",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a,c2b)]
c3b<-unique_gids %>%
  grep("NR20C3",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a,c2b,c3a)]
nrTP<-setdiff(unique_gids,unique(c(c1a,c1b,c2a,c2b,c3a,c3b)))
```

```{r}
## Format and write GEBV
predModelA %>% 
  select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(-GETGV,-contains("PEV")) %>%
  spread(Trait,GEBV) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b")) %>% 
  select(Group,GID,any_of(traits)) %>% 
  arrange(desc(Group)) %>% 
  write.csv(., file = here::here("output","GEBV_NRCRI_ModelA_2021May03.csv"), row.names = F)

## Format and write GETGV
predModelADE %>% 
  select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(GID,Trait,GETGV) %>% 
  spread(Trait,GETGV) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b")) %>% 
  select(Group,GID,any_of(traits)) %>% 
  arrange(desc(Group)) %>% 
  write.csv(., file = here::here("output","GETGV_NRCRI_ModelADE_2021May03.csv"), row.names = F)

### Make a unified "tidy" long-form: 
predModelA %>% 
  select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(-GETGV) %>% 
  full_join(predModelADE %>% 
              select(Trait,genomicPredOut) %>% 
              unnest(genomicPredOut) %>% 
              select(-varcomps) %>% 
              unnest(gblups) %>% 
              rename(GEBV_modelADE=GEBV,
                     PEV_modelADE=PEVa) %>% 
              select(-genomicPredOut)) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b")) %>% 
  relocate(Group,.before = GID) %>% 
  write.csv(., file = here::here("output","genomicPredictions_NRCRI_2021May03.csv"), row.names = F)
```

# Results

See [Results](05-Results.html): Home for plots and summary tables.
